---
title: "Real estate price prediction"
output: html_document
date: "2026-07-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
housing = read.csv("C:/Users/LENOVO/Documents/Projects/Regression Analysis/Real estate.csv",header = T)
attach(housing)
```

```{r}
# Rename the column names
names(housing) = c("ID","t.date","h.age","distance.MRT","no.stores","la.tude","lo.tude","price")
names(housing)

# Getting summary of the data set
summary(housing)
head(housing)
```

# Checking for missing values
```{r}
sum(is.na(housing))
```
#There is no missing values




# Pair each variable to get scatter plots
```{r}
pairs(housing)
```

# Get correlation matrix and round up value for 4 decimal points
```{r}
round(cor(housing),4)
```





# Variable selecting using - Adjusted R-squared value

```{r}
summary(lm(price~t.date, data = housing))$adj.r.squared
summary(lm(price~h.age, data = housing))$adj.r.squared
summary(lm(price~distance.MRT, data = housing))$adj.r.squared
summary(lm(price~no.stores, data = housing))$adj.r.squared
summary(lm(price~la.tude, data = housing))$adj.r.squared
summary(lm(price~lo.tude, data = housing))$adj.r.squared
```
#0.4524284 is the largerst value, therefore the model: price~distance.MRT 

```{r}
summary(lm(price~distance.MRT+t.date, data = housing))$adj.r.squared
summary(lm(price~distance.MRT+h.age, data = housing))$adj.r.squared
summary(lm(price~distance.MRT+no.stores, data = housing))$adj.r.squared
summary(lm(price~distance.MRT+la.tude, data = housing))$adj.r.squared
summary(lm(price~distance.MRT+lo.tude, data = housing))$adj.r.squared
```

#0.4941186 is the largerst value, therefore the model: price ~ distance.MRT + no.stores 

```{r}
summary(lm(price~distance.MRT+no.stores+t.date, data = housing))$adj.r.squared
summary(lm(price~distance.MRT+no.stores+h.age, data = housing))$adj.r.squared
summary(lm(price~distance.MRT+no.stores+la.tude, data = housing))$adj.r.squared
summary(lm(price~distance.MRT+no.stores+lo.tude, data = housing))$adj.r.squared
```

#0.5377052 is the largerst value, therefore the model: price ~ distance.MRT + no.stores + h.age

```{r}
summary(lm(price~distance.MRT+no.stores+h.age+t.date, data = housing))$adj.r.squared
summary(lm(price~distance.MRT+no.stores+h.age+la.tude, data = housing))$adj.r.squared
summary(lm(price~distance.MRT+no.stores+h.age+lo.tude, data = housing))$adj.r.squared
```

#0.5669409 is the largerst value, therefore the model: price ~ distance.MRT + no.stores + h.age + la.tude

```{r}
summary(lm(price~distance.MRT+no.stores+h.age+la.tude+t.date, data = housing))$adj.r.squared
summary(lm(price~distance.MRT+no.stores+h.age+la.tude+lo.tude, data = housing))$adj.r.squared
```

#0.5771845 is the largerst value, therefore the model: price ~ distance.MRT + no.stores + h.age + la.tude + t.date

```{r}
summary(lm(price~distance.MRT+no.stores+h.age+la.tude+t.date+lo.tude, data = housing))$adj.r.squared
```
#0.5762137 is the largerst value, therefore the model: price ~ distance.MRT + no.stores + h.age + la.tude + t.date


# Plot adjusted R-squared values
```{r}
plot(c(seq(from=1,to=6)),c(0.4524284,0.4941186,0.5377052, 0.5669409,0.5771845,0.5762137),
  xlab= "Number of variables in the  model",
  ylab= "Adjusted R-squared values")
```
#According to the plot there is no significant increase between 5th and 6th variable
#Therefore 6th variable should not include in the model
#Therefore the model is,
# price ~ price~distance.MRT+no.stores+h.age+la.tude+t.date


```{r}
lm_model = lm(price~distance.MRT+no.stores+h.age+la.tude+t.date,data = housing)
summary(lm_model)
```
#R^2 = 58.23%, Approximately 58.23% of variability in house price can be explained by this model


```{r}
predicted = predict(lm_model)
housing$predicted = predict(lm_model)
plot(housing$price, predicted,
     xlab = "Actual Price",
     ylab = "Predicted Price",
     main = "Actual vs Predicted House Prices",
     pch = 19)

abline(a = 0, b = 1, col = "red", lwd = 2)
```




```{r}
plot(lm_model)
```

# Check for normality of residuals
```{r}
hist(residuals(lm_model))
```

```{r}
shapiro.test(residuals(lm_model))
```
#The small p-value(< 2.2e-16) indicates that,we can reject the null hypothesis; 
#Residuals are not normally distributed



# Check for multicollinearity
```{r}
library(car)
round(vif(lm_model),3)
```
#Since all VIF values are below 5 , indicating no evidence of problematic multicollinearity among the predictor variable



# Check for homoscedasticity
```{r}
#install.packages("lmtest")
library(lmtest)

bptest(lm_model)
```
#p = 0.33 > 0.05 (High p-value): Fail to reject H₀; assume homoscedasticity good for standard regression



# Check for outliers
#View the Cook's Distance plot
```{r}
plot(lm_model, which = 4)

```
# Rows 149, 271 and 313 are outliers




# Confidence interval
```{r}
confint(lm_model)
```




